---
title: "HW6 document"
output: github_document
---

```{r setup}
library(tidyverse)
library(modelr)
library(viridis)
library(mgcv)
set.seed(1)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r problem 1 cleaning}
homicide_df = 
  read_csv("homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```


```{r glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

The odds ratio is 0.426, with confidence interval of 0.325 and 0.558. 


```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```
Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### Problem 2

```{r download data problem2}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


```{r Bootstrap analysis using weather data}
bootstrap_analysis = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    lm_models = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    glance_data = map(lm_models, broom::glance)
  ) |>
  select(glance_data) |>
  unnest(glance_data)

tidy_analysis = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    linear_models = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    tidy_data = map(linear_models, broom::tidy)
  ) |>
  select(tidy_data) |>
  unnest(tidy_data) |>
  select(term, estimate) |>
  pivot_wider(names_from = term, values_from = estimate) |>
  unnest()

# Bootstrap analysis with retaining 'term' and 'estimate'
tidy_analysis <- weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    linear_models = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    tidy_data = map(linear_models, broom::tidy)
  ) |>
  select(strap, tidy_data) |>
  unnest(tidy_data)

tidy_analysis_adjusted <- tidy_analysis |>
  filter(term %in% c("tmin", "prcp")) |>
  group_by(strap) |>
  summarize(
    tmin_coeff = estimate[term == "tmin"],
    prcp_coeff = estimate[term == "prcp"]
  ) |>
  mutate(
    prod_coeff = tmin_coeff * prcp_coeff,
    logarithmic_value = log(prod_coeff)
  ) |>
  filter(!is.nan(logarithmic_value) & !is.infinite(logarithmic_value))

```


```{r adjusted calculation of logarithmic values}
tidy_analysis_adjusted <- tidy_analysis |>
  filter(term %in% c("tmin", "prcp")) |>
  group_by(strap) |>
  summarize(
    prod_coeff = prod(estimate[term == "tmin"] * estimate[term == "prcp"])
  ) |>
  mutate(
    logarithmic_value = log(prod_coeff)
  ) |>
  filter(!is.nan(logarithmic_value) & !is.infinite(logarithmic_value))

```

```{r Plotting the distribution of r-squared values}
r_squared_plot = bootstrap_analysis |>
  ggplot(aes(x = r.squared)) +
  geom_density()

print(r_squared_plot)

```


```{r Plotting the distribution of the logarithmic value}
log_value_plot_adjusted = tidy_analysis_adjusted |>
  ggplot(aes(x = logarithmic_value)) +
  geom_density()

print(log_value_plot_adjusted)
```

```{r Confidence intervals for R squared}
ci_r_squared_adjusted = bootstrap_analysis |> 
  summarize(
    lower_bound_r2 = quantile(r.squared, 0.025), 
    upper_bound_r2 = quantile(r.squared, 0.975)
  )

ci_r_squared_adjusted

```

```{r Calculating the confidence interval for the log value}
ci_log_value_adjusted = tidy_analysis_adjusted |>
  summarize(
    lower_bound_log = quantile(logarithmic_value, 0.025), 
    upper_bound_log = quantile(logarithmic_value, 0.975)
  )

ci_log_value_adjusted

```
The 95% confidence interval for r square is (0.889, 0.941).
he 95% confidence interval for log (beta1*beta2) is (-4.60, -9.26).


